---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r}
#| include: false
library(tidyverse)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 300, 
  dev = "ragg_png"
)
```

# mfa-dict

This repository hosts the words we have had to add to the English
pronunciation dictionary for the Montreal Forced Aligner
([Docs][mfa-docs], [GitHub][mfa-github].


## contents

Here are the contents of the custom dictionary, as an R data-frame,
where I have named unimportant columns `"ignoreX"`.

```{r}
"english_us_arpa_extras.dict" |> 
  readr::read_tsv(
    col_names = c("word", "prob", "ignore0", "ignore1", "ignore2", "phones"),
    col_types = "cddddc"
  ) |> 
  print(n = Inf)
```

To use the dictionary, copy and paste the contents of file
`english_us_arpa_extras.dict` onto the end of the dictionary file used
by the MFA.


## updating/modifying the dictionary

I used the following to force an update to my English models.

```
mfa model download dictionary english_us_arpa --ignore_cache
mfa model download acoustic english_us_arpa --ignore_cache
```

So, for my Windows computer, the dictionary was saved to:

`C:/Users/Tristan/Documents/MFA/pretrained_models/dictionary/english_us_arpa.dict`

Then in R I can read in the MFA dictionary, our custom one, and combine them and
overwrite the MFA one.

```{r, eval = FALSE}
# Using the convention that ~ points to a user's `Documents` folder on Windows
path_mfa_dict <- "~/MFA/pretrained_models/dictionary/english_us_arpa.dict"

current <- readr::read_tsv(
  path_mfa_dict,
  col_names = FALSE,
  col_types = "cddddc"
)

new <- readr::read_tsv(
  "english_us_arpa_extras.dict",
  col_names = FALSE,
  col_types = "cddddc"
)

# Add or overwrite rows in `current` with rows from `new`
dplyr::rows_upsert(current, new, by = c("X1", "X6")) |> 
  readr::write_tsv(path_mfa_dict, col_names = FALSE)
```

### the dictionary format 

The format for the probabilistic pronunciation dictionaries is
[described here][mfa-pron-prob]. Consider the word *the*:

```
the	0.99	0.01	1.21	0.97	DH AH0
the	0.01	0.12	1.87	0.84	DH AH1
the	0.17	0.11	1.27	0.96	DH IY0
```

The docs tell us:

> The first float column is the probability of the pronunciation, the
> next float is the probability of silence following the pronunciation,
> and the final two floats are correction terms for preceding silence
> and non-silence. Given that each entry in a dictionary is independent
> and there is no way to encode information about the preceding context,
> the correction terms are calculated as how much more common was
> silence or non-silence compared to what we would expect factoring out
> the likelihood of silence from the previous word.

I am not sure what this means, but the first column seems to be the most
important (marginal) probability of pronunciation.






[mfa-docs]: https://montreal-forced-aligner.readthedocs.io/en/latest/index.html "Read The Docs page for the MFA"

[mfa-github]: https://github.com/MontrealCorpusTools/Montreal-Forced-Aligner "MFA GitHub page"

[mfa-phon-prob]: https://montreal-forced-aligner.readthedocs.io/en/latest/user_guide/dictionary.html#silence-probabilities "MFA Dictionary Format"
